\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
I would like to thank my supervisors, Ingrid and Crystal, for helping and motivating me through the process of writing this thesis. They are both highly academically gifted and genuinely nice people. It has been a pleasure to work with them, both on the thesis and the research project. I am grateful that we got to know each other on a personal level through the research trips.

I would also like to thank Adrian, Christoph, and Michael for their help and support, and for welcoming us to Germany and Braunschweig for our collaboration. The trips to Germany and our meetings in Oslo have some of the highlights during my studies, both academically and socially. I especially enjoyed the Christmas market in Braunschweig. I am very proud to have my name on a published paper alongside all these great and very clever people.

The Department of Informatics has been an excellent place to study. I have made valuable and lasting friendships here, and met some of my favourite people. Academically, I have thrived here both as a student and as a teaching assistant, largely due to the warm social environment at this department.

I am grateful to Lars for our morning coffees during the pandemic, and for helping me design my favourite symbol $\innr$. His enthusiasm inspires me.

A special thanks goes to Eirik, who has been my greatest support and friend during our studies. He has taught me a lot over the course of our various academic collaborations, patiently helping and encouraging me whenever my motivation has faltered. Our late night study sessions have been some of the best times I have had during the five years we have studied together. I am grateful to have had such a wonderful friend to write my thesis with. 

Finally, I wish to thank my parents, who have always encouraged me to pursue my interests and helped me achieve my goals. 

\begin{flushright}
   \textbf{Ida Sandberg Motzfeldt}\\
   Oslo, 2021
\end{flushright}
\end{abstract}

\renewcommand{\abstractname}{Abstract}
